Designing and modeling a browser-based distributed evolutionary computation systems

### 2015-03-29 modeling-volunteer-computing-gecco.tex
First version, now reduce

\title{Modeling browser-based distributed evolutionary computation systems}

@@

\begin{abstract}
From the era of big science we are back to the ``do it yourself'', where
you do not have any money to buy clusters or subscribe to grids but
still have algorithms that crave many computing nodes and need them to
measure scalability. Fortunately, this coincides with the era of big
data, cloud computing, and browsers that include JavaScript virtual
machines. Those are the reasons why this paper will focus on two
different aspects of volunteer or freeriding computing: first, the
pragmatic: where to find those resources, which ones can be used, what
kind of support you have to give them; and then, the theoretical: how
evolutionary algorithms can be adapted to an environment in which nodes
come and go, have different computing capabilities and operate in
complete asynchrony of each other. We will examine the setup needed to
create a very simple distributed evolutionary algorithm using
JavaScript and then find a model of how users react to it by
collecting data from several experiments featuring different classical
benchmark functions. 
\end{abstract}

@@

The world has computational resources in spades. Most of them do not
belong to you or your lab. That does not mean you cannot use them. The
problem is how.

Most theory in parallel computing has been devoted to predict and
optimize the performance in systems where the number of nodes, their
connections, and the time every one is dedicating to the computation is
known in advance. However, even if Big Science is not really over and
it is slated to come back, the era of
Citizen Science already started a few years ago (with SETI@home \cite{david-seti:home} and then BOINC \cite{boinc_grid04}) and it
offers a vast amount of computational resources to be used, if only
you know how to attract them. But there is a challenge: knowing, or at least having a
ballpark, of how your algorithm is going to perform in this uncertain
environment, where none of the factors is known: neither the number of
nodes, through how they are connected, to how long are they going to
be focused on doing what you want them to. That is why some effort is
needed to first understand the dynamics of the decision to participate
in an experiment that requires you to click on a link and then stay
for a while looking at the screen (or just leave it there running).

>>>>

\title{Designing and modeling a browser-based distributed evolutionary computation systems}

@@

\begin{abstract}
Browsers have scaled from simple page-rendering engines to operating
systems that include most services the lower OS layer has, with the
added facility that applications can be run by just visiting a web
page. In this paper we will describe the front and back end of a
distributed evolutionary computation system that uses the browser's
capabilities of running programs written in JavaScript. We will focus on two
different aspects of volunteer computing: first, the
pragmatic: where to find those resources, which ones can be used, what
kind of support you have to give them; and then, the theoretical: how
evolutionary algorithms can be adapted to an environment in which nodes
come and go, have different computing capabilities and operate in
complete asynchrony of each other. We will examine the setup needed to
create a  simple distributed evolutionary algorithm using
JavaScript and then find a model of how users react to it by
collecting data from several experiments featuring different classical
benchmark functions. 
\end{abstract}

@@

The world has computational resources in spades, with every person
wielding one or more computational devices. The issue is how to make
them allocate even a small part of those resources to your particular
experiment.

Most theory and experimentation in parallel computing has been devoted to predict and
optimize the performance in systems where the number of nodes, their
connections, and the time every one is dedicating to the computation is
known in advance as it often did in the Big Science era, where labs
could easily buy 16-node clusters or subscribe to the resources of a
big grid. Even today, you can rent instances of virtual machines in
the cloud relatively cheaply and use them for evolutionary computation
experiments \cite{sherry2012flex}. Nothing beats free, which is one of
the reasons why the era of
Citizen Science already started a few years ago (with SETI@home \cite{david-seti:home} and then BOINC \cite{boinc_grid04}) and it
offers a vast amount of computational resources to be used, if only
you know how to attract them. But there is a challenge: knowing, or at least having a
ballpark, of how your algorithm is going to perform in this uncertain
environment, where none of the factors is known: neither the number of
nodes, through how they are connected, to how long are they going to
be focused on doing what you want them to. That is why some effort is
needed to first understand the dynamics of the decision to participate
in an experiment that requires you to click on a link and then stay
for a while looking at the screen (or just leave it there running).
